Automated topic naming to support cross-project analysis of software maintenance activities

### 2011-01-26 msr/msr2011.tex
migod edits II

In order to automatically label each topic with an NFR from ISO9126, we associate each NFR with a list
of keywords (`word-lists', in our parlance). These word-lists were determined a priori and were not extracted from the projects themselves.
We intersected the words of the topics and the words of our word-lists.
We `labeled' a topic if any of its words matched any of the word-list's words.
A topic could match more than one NFR.
We used several different sets of word-lists for comparison, which we refer to as \textsf{exp1, exp2, exp3} in the text which follows.

@@

Our second word-list, \textsf{exp2}, uses the ISO9126 taxonomy described above (Section \ref{sec:wordlist}) to seed the word-lists.
%We use these qualities later on as classes in supervised labelling. NE confusing the issue here, I think.
The terms from ISO9126 may not capture all words occurring in the topics that are nonetheless associated with one of the NFRs. 
For example, the term ``redundancy'' is one most would agree is relevant to discussion of reliability, but is not in the standard. 
We therefore took the NFRs from the ISO9126 and added to them.

@@

Using our three word-lists (\textsf{exp1}, \textsf{exp2}, \textsf{exp3}), we labeled our topics with an NFR where there was a match between a word in the list and the same word somewhere in the distribution of words that constitute the topic.
A \emph{named topic} is a topic with a matching NFR label. Unnamed topics occur where this no such match, which may indicate either a lack of precision, or simply that this topic is not associated with non-functional requirements.
All experiments were run on MaxDB 7.500 and MySQL 3.23 data. LDA
extracted 20 topics per period for each.
% Is this good enough
This labelling is unsupervised because the corpus is not derived from 
the project being labelled, and we did not label the project's topics
ourselves for a training set. The theory behind this technique is that
because most software often addresses similar issues, we can use the
domain knowledge of software to label relevant topics.
%\todo[inline]{Neil does this explain the unsupervised aspect well enough?}

@@

For \textsf{exp1}, our best performing labels, those with the most topics, were \emph{correctness} (182 topics) and \emph{testability} (121). 
We did not get good results for \emph{usability} or \emph{accuracy}, which were associated with fewer than ten topics. 
We also looked for correlation between our labels: Excluding double matches (self-correlation), our highest co-occurring terms were \emph{verifiability} with \emph{traceability}, and \emph{testability} with \emph{correctness} (76 and 62 matches, respectively).
% See the following sections for our error analysis.

For \textsf{exp2}, there are more unnamed topics than \textsf{exp1}. 
Only \emph{reliability} produces a lot of matches, mostly with the word `error'. 
Co-occurrence results were poor. This suggests our word lists were overly restrictive.

For \textsf{exp3}, we generally labelled more topics. 
As we mentioned, the word-lists are quite broad, so there are likely to be false-positives (discussed below). 
We found a high of 265 topics for \emph{usability}, with a low of 44 topics for \emph{maintainability}. 
Common co-occurrences were \emph{reliability} with \emph{usability}, \emph{efficiency} with \emph{reliability}, and \emph{efficiency} with \emph{usability} (200, 190, and 150 topics in common, respectively).

@@

%EXP1 could not be evaluated properly!
Because our ground truth annotations were only relevant to ISO-9126,
we estimate that \textsf{exp1} had poor
performance via the overlap between ISO-9126 and the Kayed ontology.
For \textsf{exp1} the F-measures for MaxDB were from $0$ to $0.18$ with an average
of $0.03$, and for MySQL were from $0$ to $0.16$ with an average of
$0.05$.
% 0.177215189873 / 6
% .02953586497883333333
% (0.0298507462687 + 0.0661157024793 + 0.0634920634921 + 0 + 0.115384615385) / 6
% .04580718793751666666

@@

Yes. MySQL 3.23 had proportionally more
topics labelled \emph{functionality}, while MaxDB had proportionally more
\emph{efficiency} related topics. MaxDB was a very mature release `donated' to the open-source community. 
MySQL, on the other hand, was in its relative infancy. 
Security problems were more common (security is a component of `functionality' in the ISO 9126 model). 
In both cases \emph{portability} was a constant maintenance concern and was prevalent throughout the entire lifetime of the projects.

@@

%XXX can just remove this
%\subsection{Effectiveness}
\subsection{Annotation observations}
We found many topics that were not non-functional requirements (NFRs) but were often related to them. 
For instance, concurrency was mentioned often in the commit logs and was related to \emph{correctness} and \emph{reliability}, likely because concurrent code is prone to bugs such as race conditions. %are 
% often  was troublesome. 
\emph{Configuration management} and \emph{source control} related changes appeared often; % and sometimes there were topics dedicated to configuration management. 
these kinds of changes are slightly related to \emph{maintainability}. 
A non-functional change that was not quality-related was licensing and copyright; many changes were simply to do with updating copyrights or ensuring copyright or license headers were applied to files. In these cases we assigned the \emph{None} class.

@@

\section{Conclusions}
This paper presented a cross-project data mining technique, \emph{labelled topic extraction}. The technique leveraged a non-functional requirements (NFRs) taxonomy to mine and label latent software development topics. Previous topic analysis research produced topics that needed to be manually labelled, but topics without labels are difficult to interpret, and manually labelling them is costly. Furthermore, previous work was project-specific and incomparable. 
To improve this, we leveraged software engineering standards, specifically the ISO quality taxonomy, to produce a method of supervised and unsupervised topic labelling.
Since the word-list technique is not project specific, it can be used to compare multiple projects.%topics and artifacts from project histories.
% Our contributions include:
% \begin{itemize}
% \item We provided an unsupervised method of topic labelling (word-lists);
% %\item Our unsupervised method can be applied to projects without training;
% \item We demonstrated a supervised method of labelling topics by a single NFRs (machine learning);
% \item We demonstrated a supervised method of labelling topics by multiple NFRs (multi-label machine learning);
% \item We provided a method of cross-project analysis via topic labelling;
% \item We demonstrated a method of visualizing these topics across time, which we used to analyze maintenance activities.
% \end{itemize}

>>>>

In order to automatically label each topic with 
one of the six high-level NFRs,
%an NFR from ISO9126, 
we associate each NFR with a list
of keywords (`word-lists', in our parlance). These word-lists were determined a priori and were not extracted from the projects themselves.
We intersected the words of the topics and the words of our word-lists.
We ``labeled'' a topic if any of its words matched any of the word-list's words.
A topic could match more than one NFR.
We used several different sets of word-lists for comparison, which we
refer to as \textsf{exp1}, \textsf{exp2}, and \textsf{exp3} in the text which follows.

@@

Our second word-list, \textsf{exp2}, uses the ISO9126 taxonomy described above (Section \ref{sec:wordlist}) to seed the word-lists.
%We use these qualities later on as classes in supervised labelling. NE confusing the issue here, I think.
The terms from ISO9126 may not capture all words occurring in the topics that are nonetheless associated with one of the NFRs. 
For example, the term ``redundancy'' is one 
we considered to be
relevant to discussion of reliability, but is not in the standard. 
We therefore took the NFRs from the ISO9126 and added to them.

@@

Using our three word-lists (\textsf{exp1}, \textsf{exp2}, \textsf{exp3}), we labeled our topics with an NFR where there was a match between a word in the list and the same word somewhere in the distribution of words that constitute the topic.
A \emph{named topic} is a topic with a matching NFR label. Unnamed topics occur where this no such match, which may indicate either a lack of precision, or simply that this topic is not associated with non-functional requirements.
All experiments were run on MaxDB 7.500 and MySQL 3.23 data. LDA
extracted 20 topics per period for each.
% Is this good enough
This labelling is unsupervised because the corpus is not derived from 
the project being labelled, and we did not label the project's topics
ourselves for a training set. The motivation behind this technique is that
because most software often addresses similar issues, we can use the
domain knowledge of software to label relevant topics.
%\todo[inline]{Neil does this explain the unsupervised aspect well enough?}

@@

For \textsf{exp1}, our best performing 
labels --- that is, those with the most topics --- were
%labels, those with the most topics, 
\emph{correctness} (182 topics) and \emph{testability} (121). 
We did not get good results for \emph{usability} or \emph{accuracy}, which were associated with fewer than ten topics. 
We also looked for correlation between our labels: Excluding double matches (self-correlation), our highest co-occurring terms were \emph{verifiability} with \emph{traceability}, and \emph{testability} with \emph{correctness} (76 and 62 matches, respectively).
% See the following sections for our error analysis.

For \textsf{exp2}, there are more unnamed topics than \textsf{exp1}. 
Only \emph{reliability} produces a lot of matches, mostly with the word ``error''. 
Co-occurrence results were poor. This suggests our word lists were overly restrictive.

For \textsf{exp3}, we generally labelled more topics. 
As we mentioned, the word-lists are broad, so there are likely to be false-positives (discussed below). 
We found a high of 265 topics for \emph{usability}, with a low of 44 topics for \emph{maintainability}. 
Common co-occurrences were \emph{reliability} with \emph{usability}, \emph{efficiency} with \emph{reliability}, and \emph{efficiency} with \emph{usability} (200, 190, and 150 topics in common, respectively).

@@

%EXP1 could not be evaluated properly!
Because our ground truth annotations were relevant only to ISO9126,
we estimate that \textsf{exp1} had poor
performance via the overlap between ISO-9126 and the Kayed ontology.
For \textsf{exp1} the F-measures for MaxDB were from $0$ to $0.18$ with an average
of $0.03$, and for MySQL were from $0$ to $0.16$ with an average of
$0.05$.
% 0.177215189873 / 6
% .02953586497883333333
% (0.0298507462687 + 0.0661157024793 + 0.0634920634921 + 0 + 0.115384615385) / 6
% .04580718793751666666

@@

Yes. MySQL 3.23 had proportionally more
topics labelled \emph{functionality}, while MaxDB had proportionally more
\emph{efficiency} related topics. MaxDB was a very mature release ``donated'' to the open-source community. 
MySQL, on the other hand, was in its relative infancy. 
Security problems were more common (security is a component of `functionality' in the ISO 9126 model). 
In both cases \emph{portability} was a constant maintenance concern and was prevalent throughout the entire lifetime of the projects.

@@

%XXX can just remove this
%\subsection{Effectiveness}
\subsection{Annotation observations}
We found many topics that were not non-functional requirements (NFRs) but were often related to them. 
For instance, concurrency was mentioned often in the commit logs and
was related to \emph{correctness} and \emph{reliability}, possibly because concurrent code is prone to bugs such as race conditions. %are 
% often  was troublesome. 
\emph{Configuration management} and \emph{source control} related changes appeared often; % and sometimes there were topics dedicated to configuration management. 
these kinds of changes are slightly related to \emph{maintainability}. 
A non-functional change that was not quality-related was licensing and copyright; many changes were simply to do with updating copyrights or ensuring copyright or license headers were applied to files. In these cases we assigned the \emph{None} class.

@@

\section{Conclusions}
This paper presented a cross-project data mining technique,
\emph{labelled topic extraction}. The technique leveraged a
non-functional requirements (NFRs) taxonomy to mine and label latent
software development topics. Previous topic analysis research produced
topics that needed to be manually labelled, but topics without labels
are difficult to interpret, and manually labelling them is
costly. Furthermore, previous work was project specific and incomparable. 
To improve this, we leveraged software engineering standards, specifically the ISO quality taxonomy, to produce a method of supervised and unsupervised topic labelling.
Since the word-list technique is not project specific, it can be used to compare multiple projects.%topics and artifacts from project histories.
% Our contributions include:
% \begin{itemize}
% \item We provided an unsupervised method of topic labelling (word-lists);
% %\item Our unsupervised method can be applied to projects without training;
% \item We demonstrated a supervised method of labelling topics by a single NFRs (machine learning);
% \item We demonstrated a supervised method of labelling topics by multiple NFRs (multi-label machine learning);
% \item We provided a method of cross-project analysis via topic labelling;
% \item We demonstrated a method of visualizing these topics across time, which we used to analyze maintenance activities.
% \end{itemize}
